Embedding Intuitionistic into Classical Logic

### 2023-03-18 LPAR/LPAR-EIICL.tex
Proof A.2

\begin{document}
	%
	\title{Embedding Intuitionistic into Classical Logic}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	\author{Alexander Pluska \and Florian Zuleger}
	%
	\authorrunning{A. Pluska and F. Zuleger}
	\titlerunning{Embedding Intuitionistic into Classical Logic}
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{TU Wien, Vienna, Austria}
	%
	\maketitle
\begin{abstract}
The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding of classical into intuitionistic logic.
Curiously, the reverse direction has not been covered in literature.
Utilizing a normal form for intuitionistic logic~\cite{claessen2015sat}, we establish a small model property for intuitionistic propositional logic.
We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified boolean formulae.
Next, we transfer the developed techniques to the first order case and provide an embedding of intuitionistic first-order logic into classical first-order-logic.
Our goal here is an encoding that facilitates the use of state-of-the-art provers for classical first-order logic for determining intuitionistic validity.
In an experimental evaluation, we show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
We further note that our constructions support the transfer of counter-models to validity, which is a desired feature in model checking applications.
\end{abstract}

@@

With this work we want to leverage the progress in classical reasoning for intuitionistic logic.
To this end, we study embeddings of intuitionistic into classical logic and propose the following methodology:
\begin{itemize}
	\item Give for each $\varphi$ a formula $\varphi^\#$, which is classically valid iff $\varphi$ is intuitionistically valid.
	\item Determine the classical validity of $\varphi^\#$ using a prover for classical logic.
	\item Translate the classical proof (resp. counter-model) of $\varphi^\#$ given by the prover to a corresponding intuitionistic proof (resp. counter-model) of $\varphi$.
\end{itemize}
\subsection*{Contributions}
The following main contributions will be discussed in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}:
\begin{itemize}
	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulae.
	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding of intuitionistic into classical first-order logic.
	\item We provide an experimental evaluation on the use of first-order provers for classical logic for determining intuitionistic validity.
 We show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
\end{itemize}

\subsection*{Related work}
The famous double-negation translation establishes
an embedding of classical into intuitionistic logic, going back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
In the propositional case the embedding is particularly simple:
A formula $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
Intuitively, the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
For embedding intuitionistic into classical logic, we need to do the opposite:
We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e., if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
This necessitates the introduction of new propositional variables in our procedure, which marks a big difference to the double negation translation.

@@

We finally mention a long-standing line of research that has investigated when classical validity implies intuitionistic validity.
This most famously holds for $\Pi_2$-formulae in Peano Arithmetic~\cite{friedman1978classically}, but also other conditions are known~\cite{schwichtenberg}.
In contrast, we propose a translation of an intuitionistic formula to an equi-valid classical formula.

@@

The details can be found in Section~\ref{sec:ipc-to-cpc}.
Building on the same idea, we can also give a quantified boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers. We obtain the following result:

@@

\begin{definition}
	Let $\varphi$ be a propositional formula.
 Define $\varphi^{b}$ inductively:
	\vspace*{-.1cm}
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A^{b} = A(b)$ for every propositional variable $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^b$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
	\vspace*{-.1cm}
	\[K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
	\vspace*{-.4cm}
	with e.g.
	\vspace*{-.2cm}
	\begin{align*}
		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$\varphi^{C} = K(\varphi)\to \varphi^{b}$
	where $b$ is a new free variable.
\end{definition}

@@

Since $|\Lambda|\in O(2^{|\mathcal X|\log |\mathcal X|})$ a simple counting argument shows that $|\mathcal P^\#|\in\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log |\mathcal X|})$, $|\mathcal R^\#|\in\mathcal O(|\mathcal R|\cdot2^{|\mathcal X|\log |\mathcal X|})$ and $|\mathcal X|\in \mathcal O(|\mathcal X|\cdot 2^{|\mathcal X|\log|\mathcal X|})$. This gives us a bound on the size of $\varphi^#$.

\begin{lemma}
	The size of $\varphi^\#$ is in $\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
\end{corollary}

@@

\begin{definition}
For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the $n$-th position of a node on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$.
Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula $\psi$ at the $n$-th position of the sequence, and it is at a position that is $0$ in all $\vec X^i$ with $i < n$ (i.e., it did not occur in the sequence previously).
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulae in $\mathcal R$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
	\end{itemize}
	Define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < |\mathcal{X}|$, as well as the special cases
	$$\varphi^Q_{|\mathcal{X}|} = \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
	for leaves, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulae are:
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j \in \mathcal R\right\}\\
		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
	\end{align*}
\end{definition}

@@

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
\todo{genau genommen müsstest du noch sagen, dass $b$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$.
		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
		\item $(\forall x\varphi)^b = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
	\end{itemize}
	where $w$ is some new bound variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:\vspace*{-.76cm}
	\begin{align*}
		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) = &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) = &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) = &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$ where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

As in the previous section we can define a tree counter-model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. W.l.o.g. suppose $\{m\in M\:|\:\text{World}(m)\}$ has a minimal element, say $w_0$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the union of $M$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, w_0)$ and $\vec E^I(\vec m_i, f_{\psi_i}(\vec m_i, f_{\psi_{i-1}}^I(\vec m_{i-1}, \dots f_{\psi_1}^I(\vec m_1, w_0)\dots)))$ for $i > 0$ hold.
		\item $s^{I_T} = w_0$, $\vec r^{I_T} = \vec r^I$.
		\item We interpret $\preceq$ trivially on $M$ and as the prefix-order elsewhere.
		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, f_{\psi_n}^I(\vec m_n, \dots f_{\psi_1}^I(\vec m_1, w_0)\dots))$.
		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi)$ and arbitrary else.
		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}^I(\vec m_n, \dots(f_{\psi_1}^I(\vec m_1, s^I))\dots))$$ and arbitrary otherwise, for function symbols $f^{I_T}(\vec m) = f^I(\vec m)$ and arbitrary else.
	\end{itemize}
\end{definition}

@@

\begin{figure}
	\centering
	\begin{tabular}{c|c|lr|lr|lr|lr|c}
		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Embedding (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
	\end{tabular}
	\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
	\label{fig:b}
\end{figure}

@@

>>>>

\begin{document}
	%
	\title{Embedding Intuitionistic into Classical Logic}
	%
	%\titlerunning{Abbreviated paper title}
	% If the paper title is too long for the running head, you can set
	% an abbreviated paper title here
	%
	\author{Alexander Pluska \and Florian Zuleger}
	%
	\authorrunning{A. Pluska and F. Zuleger}
	\titlerunning{Embedding Intuitionistic into Classical Logic}
	% First names are abbreviated in the running head.
	% If there are more than two authors, 'et al.' is used.
	%
	\institute{TU Wien, Vienna, Austria}
	%
	\maketitle
\begin{abstract}
The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding of classical into intuitionistic logic.
Curiously, the reverse direction has not been covered in literature.
Utilizing a normal form for intuitionistic logic~\cite{claessen2015sat}, we establish a small model property for intuitionistic propositional logic.
We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified Boolean formulae.
Next, we transfer the developed techniques to the first order case and provide an embedding of intuitionistic first-order logic into classical first-order-logic.
Our goal here is an encoding that facilitates the use of state-of-the-art provers for classical first-order logic for determining intuitionistic validity.
In an experimental evaluation, we show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
We further note that our constructions support the transfer of counter-models to validity, which is a desired feature in model checking applications.
\end{abstract}

@@

With this work we want to leverage the progress in classical reasoning for intuitionistic logic.
To this end, we study embeddings of intuitionistic into classical logic and propose the following methodology:
\begin{itemize}
	\item Give for each $\varphi$ a formula $\varphi^\#$, which is classically valid iff $\varphi$ is intuitionistically valid.
	\item Determine the classical validity of $\varphi^\#$ using a prover for classical logic.
	\item Translate the classical proof (resp. counter-model) of $\varphi^\#$ given by the prover to a corresponding intuitionistic proof (resp. counter-model) of $\varphi$.
\end{itemize}
\subsection*{Contributions}
The following main contributions will be discussed in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}:
\begin{itemize}
	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified Boolean formulae.
	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding of intuitionistic into classical first-order logic.
	\item We provide an experimental evaluation on the use of first-order provers for classical logic for determining intuitionistic validity.
 We show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
\end{itemize}

\subsection*{Related work}
The famous double-negation translation establishes
an embedding of classical into intuitionistic logic, going back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to Gödel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
In the propositional case the embedding is particularly simple:
A formula $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
Intuitively, the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
For embedding intuitionistic into classical logic, we need to do the opposite:
We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e., if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
This necessitates the introduction of new propositional variables in our procedure, which marks a big difference to the double negation translation.

@@

We finally mention a long-standing line of research that has investigated when classical validity implies intuitionistic validity.
This most famously holds for $\Pi_2$-formulae in Peano Arithmetic~\cite{friedman1978classically}, but also other conditions are known~\cite{schwichtenberg}.
In contrast, we propose a translation of an intuitionistic formula to an equivalid classical formula.

@@

The details can be found in Section~\ref{sec:ipc-to-cpc}.
Building on the same idea, we can also give a quantified Boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers. We obtain the following result:

@@

\begin{definition}
	Let $\varphi$ be a propositional formula.
 Define $\varphi^{b}$ inductively:
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A^{b} = A(b)$ for every propositional variable $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^b$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
	\[K(\varphi) := \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
	with e.g.
	\begin{align*}
		\text{\normalfont PartialOrder}(\preceq) :=&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{\normalfont Persistent}(u, w):=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	\[\varphi^{C} := K(\varphi)\to \varphi^{b}\]
	where $b$ is a new free variable.
\end{definition}

@@

Since $|\Lambda|\in O(2^{|\mathcal X|\log |\mathcal X|})$ a simple counting argument shows that $|\mathcal P^\#|\in\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log |\mathcal X|})$, $|\mathcal R^\#|\in\mathcal O(|\mathcal R|\cdot2^{|\mathcal X|\log |\mathcal X|})$ and $|\mathcal X|\in \mathcal O(|\mathcal X|\cdot 2^{|\mathcal X|\log|\mathcal X|})$. This gives us a bound on the size of $\varphi^\#$.

\begin{lemma}
	The size of $\varphi^\#$ is in $\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
\end{lemma}

@@

\begin{definition}
For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the $n$-th position of a node on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$.
Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula $\psi$ at the $n$-th position of the sequence, and it is at a position that is $0$ in all $\vec X^i$ with $i < n$ (i.e., it did not occur in the sequence previously).
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulae in $\mathcal R$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
	\end{itemize}
	Define$$\varphi^Q_{i} := \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < |\mathcal{X}|$, as well as the special cases
	$$\varphi^Q_{|\mathcal{X}|} := \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
	for leaves, and $$\varphi^Q = \varphi_0^Q := \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulae are:
	\begin{align*}
		\text{\normalfont Valid}(n) := &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
		\text{\normalfont Persistent}(n) := & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
		\text{\normalfont Sat}_{\mathcal R}(n) := &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j \in \mathcal R\right\}\\
		\text{\normalfont Sat}_{\mathcal X}(n) := & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
	\end{align*}
\end{definition}

@@

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$.
		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
		\item $(\forall x\varphi)^b = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
	\end{itemize}
	where $w$ is some new bound variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) := \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:
	\begin{align*}
		\text{PartialOrder}(\preceq) := &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) := &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) := &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) := &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) := &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	\[\varphi^C:= K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b\] where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

As in the previous section we can define a tree counter-model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. Similarly to the propositional case for $\psi\in\mathcal X$ define \[g_\psi :M^n\to M, (\vec z, m) \mapsto
	\begin{cases} 
	m,&\text{if }m\not\preceq f_\psi(\vec z, m)\\
	f_{[\psi]}^I(\vec z, m)&\text{otherwise}
	\end{cases}
	\]where $[\psi]$ denotes the corresponding element of $\mathcal X^{bS}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the union of $\{m\in M\:|\:\text{World}^I(m)\wedge s^I\preceq^I m\}$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, w_0)$ and $\vec E^I(\vec m_i, g_{\psi_i}(\vec m_i, g_{\psi_{i-1}}(\vec m_{i-1}, \dots g_{\psi_1}(\vec m_1, w_0)\dots)))$ for $i > 0$ hold.
		\item $s^{I_T} = s^I$, $\vec r^{I_T} = \vec r^I$.
		\item We interpret $\preceq$ trivially on $M$ and as the prefix-order elsewhere.
		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, g_{\psi_n}(\vec m_n, \dots g_{\psi_1}(\vec m_1, w_0)\dots))$.
		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi)$ and arbitrary else.
		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, g_{\psi_n}(\vec m_n, \dots(g_{\psi_1}(\vec m_1, s^I))\dots))$$ and arbitrary otherwise, for function symbols $f^{I_T}(\vec m) = f^I(\vec m)$ and arbitrary else.
	\end{itemize}
\end{definition}

@@

\begin{figure}[h]
	\centering
	\begin{tabular}{c|c|lr|lr|lr|lr|c}
		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Embedding (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
	\end{tabular}
	\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
	\label{fig:b}
\end{figure}

@@

\newpage
